Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-07 ICPP-19/communications.tex
rearranging stuff

	%%\begin{CCSXML}
	%%	<ccs2012>
	%%	<concept>
	%%	<concept_id>10010520.10010553.10010562</concept_id>
	%%	<concept_desc>Computer systems organization~Embedded systems</concept_desc>
	%%	<concept_significance>500</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10010520.10010575.10010755</concept_id>
	%%	<concept_desc>Computer systems organization~Redundancy</concept_desc>
	%%	<concept_significance>300</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10010520.10010553.10010554</concept_id>
	%%	<concept_desc>Computer systems organization~Robotics</concept_desc>
	%%	<concept_significance>100</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10003033.10003083.10003095</concept_id>
	%%	<concept_desc>Networks~Network reliability</concept_desc>
	%%	<concept_significance>100</concept_significance>
	%%	</concept>
	%%	</ccs2012>
	%%\end{CCSXML}
	\ccsdesc[100]{Computer systems organization~Embedded systems1}
	\ccsdesc[100]{Computer systems organization~Redundancy}
	\ccsdesc{Computer systems organization~Robotics}
	\ccsdesc[100]{Networks~Network reliability}

@@

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges come into picture. First, every architecture provides its own efficacy and interface. Therefore, a steep learning curve is required for programmers to take good utilization of all resources. Second, scheduling is a well known NP-Complete optimization problem, and hybrid and distributed resources make this problem harder (we refer ~\cite{webpagescheduling} for a survey on the complexity of scheduling problems and ~\cite{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses, parallel resources, it is hard to obtain a precise model based on prediction of computation and communication timings. Fourth, number of architectures has increased drastically in recent years, therefore it is almost impossible to develop hand tune optimized code for all these architectures. All these observations led to the development of different task based runtime systems. Among several runtimes, we may cite QUARK~\cite{YarKhan:2011:Quark:Manual} and PaRSEC~\cite{parsec} from from ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} form INRIA Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAPPI~\cite{kaapi} from INRIA Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information such as distance to exit (last) node as a priority or affinity of computation towards resources is also provided along with the DAG. Then runtime manages scheduling of computations and communications, data transfers among different memories, computation-communication overlap, load balance in the best possible way.

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities, based on the task graph structure, with computations, which runtime uses to make the execution efficient. In some applications, such as sparse linear algebra kernels, where programmers do not know the precise structure of the tasks graphs in advance,  tasks are added recursively based on certain sentinel constraints. For such applications, runtime sees a set of independent ready tasks and schedule them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations such that overlap between computations and communications can be maximised. This is the main topic of this article. We propose different strategies for runtimes such that it can maximize the overlap of computation and communication for kernels running on two different memory nodes. We prove that one of our proposed strategies is an optimal one when we do not have any restriction on the capacity of memory node. We also prove that order of communications on two memory nodes  with the objective of minimizing the makespan is  NP-Complete. We evaluate our strategies on a cluster of homogeneous nodes. But our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper.

@@

	There is also a numerous study pertaining to the scheduling with limited memory and storage, since the work of Sethi and Ulman, register allocation for arithmetic expressions~\cite{Sethi:1970:GOC:321607.321620}.  Sarkar et. al  worked on the scheduling of graphs of smaller-grain tasks with limited memory, where each task requires homogeneous data size ~\cite{vsarkar-pact}. The same work has been extend by Marchal et. al for task graphs where memory requirement of each task is highly irregular~\cite{loris-ipdps18}.

@@

	In order to prove the optimality of makespan for the schedule obtained by the above algorithm, we need the following lemma.

@@

		\begin{eqnarray*}
			\scomp(A) & = & max(t_1 + CM_A, t_2) \\
			\scomp(B) & = & max(\scomp(A) + CP_A,  t_1 + CM_A + CM_B)
		\end{eqnarray*}

		Completion time of $B$ in original Schedule = $\scomm(B) + CP_B$

		\begin{eqnarray*}
			\text{Completion time of $B$ in swapped Schedule, } \scomp '(B) & = & max(t_1 + CM_B, t_2)\\
			\text{Completion time of $A$ in swapped Schedule, } \scomp '(A) & = & max(\scomp'(B) + CP_B, t_1 + CM_B + CM_A)
		\end{eqnarray*}

@@

		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(\scomp(A) + CP_A + CP_B,  t_1 + CM_A + CM_B + CP_B)\\
			& = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B) 
		\end{eqnarray*}

		Case I: $CP_A \ge CM_A, CP_B \ge CM_B, CM_A \le CM_B$
		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B)\\
			& \le& max(t_1 + CM_A + CP_B, t_2  + CP_B, t_1  + CM_B + CP_B) + CP_A\\
			& = & max( t_1  + CM_B + CP_B, t_2  + CP_B) + CP_A \\
			& = & max(max(t_1 + CM_B, t_2) + CP_B, t_1 + CM_A + CM_B) + CP_A\\
			& = & max(\scomp'(B) + CP_B, t_1 + CM_A + CM_B) + CP_A \\
			& = & \scomp'(A) + CP_A
		\end{eqnarray*}
		Case II: $CP_A < CM_A, CP_B < CM_B, CP_A \ge CP_B$
		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B)\\
			& \le & max(t_1 + CM_A + CP_B, t_2+CP_B, t_1 + CM_A + CM_B) + CP_A\\
			& = & max(t_1 + CM_A + CM_B, t_2+ CP_B) + CP_A \\
			& = & max(t_1 + CM_B + CP_B, t_2 + CP_B, t_1 + CM(A) + CM_B) + CP_A \\
			& = & max(max(t_1 + CM_B, t_2) + CP_B, t_1 + CM_A + CM_B) + CP_A \\
			& = & max(\scomp'(B) + CP_B, t_1 + CM_A +CM_B) + CP_A \\
			& = & \scomp'(A) + CP_A
		\end{eqnarray*}
		Case III: $CP_A \ge CM_A, CP_B < CM_B$
		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B)\\
			& \le & max(t_1 + CM_A + CP_B, t_2 + CP_B, t_1 + CM_B + CP_B) + CP_A\\
			& \le & max(t_1 + CM_A + CM_B, t_2 + CP_B, t_1 + CM_B + CP_B) + CP_A \\
			& = & max(max(t_1 + CM_B, t_2) + CP_B, t_1 + CM_A + CM_B) + CP_A \\
			& = & max(\scomp'(B) + CP_B, t_1 + CM_A +CM_B) + CP_A \\
			& = & \scomp'(A) + CP_A
		\end{eqnarray*}
	\end{proof}

@@

		Let the final swapped schedule is $S_{final}$. From Lemma~\ref{lemma:swappingOfTasks}, $makespan(S_{final}) \ge makesapn(S)$. From the construction, $makespan(S_{final}) \le makesapn(O)$. As $O$ is an optimal schedule, hence $makespan(S) = makesapn(O)$. This completes our proof.

@@

		Conversely, if such a partition of $A$ is achievable, then we can construct a feasible schedule $S$ 
		without idle times by the pattern depicted in Fig~\ref{fig:firstSegment}. We have shown that $3PAR$ 
		problem reduces to $DT$, and hence problem $DT$ is NP-complete.
		\todo[inline]{Skumar: Increase width of both channels of Figure~\ref{fig:firstSegment}.}

@@

	The proof of the above theorem is inspired from work by [Papadimitriou et. al], which applies a similar 
	technique for 2-machine flowshop problem, where each machine can hold fixed amount of jobs. The 
	main difference with the our approach is that they consider unit buffer requirement on 1st machine 
	after job completion. On the contrary, our approach is designed where there is not any restriction on 
	requested input data and memory is acquired before starting the data transfer on communication 
	resource.

@@

	Fig~\ref{fig:staticOrderExample} shows schedules for all proposed heuristics of this class for task set of Table~\ref{tab:staticOrderExample}.
	\subsection{Dynamic Selection}
	%%%%string alg_name="largest_communication\_task\_respects\_memory\_restriction";
	%%%%string alg_name="smallest_communication\_task\_respects\_memory\_restriction";
	%%%%string alg_name="maximum_accelerated\_task\_respects\_memory\_restriction";

@@

	Fig~\ref{fig:dynamicSelectionExample} shows schedules for all proposed heuristics of this class for task set of Table~\ref{tab:dynamicSelectionExample}.

@@

	Fig~\ref{fig:staticOrderDynamicCorrectionsExample} shows schedules for all proposed heuristics of this class for task set of Table~\ref{tab:staticOrderDynamicCorrectionsExample}.

@@

@@

	We refer ~\cite{Gilmore-Gomory:1964} for the detailed algorithm and ~\cite{gitworkrepo} for our implementation. This algorithm does not take memory capacity into consideration and only provides the sequence of processing. We applied the same sequence with different memory capacity restrictions for our experiments and call this heuristic as \textit{Gilmore-Gomory}.

	%%	\todo[inline]{may convert 10 subplots to 8 in all performance diagrams}
	\todo[inline]{increase font size in all plots, may be free scale-y}
	\todo[inline]{SKumar: decide metrics: ratio to optimal, percentage overlap, 100 - percentage non overlap}

@@

	Figures ~\ref{fig:ratio_to_optimal_hf} and ~\ref{fig:ratio_to_optimal_ccsd} depict the distribution of the performance  of each heuristic for all considered memory capacities, where plots are categorized by memory capacities. For each memory capacity and each heuristic, the box on the plot displays the median, first and last quartile, and the whiskers indicate minimum and maximum values, with outliers are shown by black dots.
	\subsection{HF Performance}	
	\begin{figure*}[htb]
		\includegraphics[scale=0.25]{./results/plots/ratio_to_optimal_selected_hf.pdf}
		\caption{Comparison of different heuristics for HF.}
		\label{fig:ratio_to_optimal_hf}
	\end{figure*}
	As indicated above, HF tasks are highly homogeneous, this is also noticeable in Figure~\ref{fig:ratio_to_optimal_hf}. All heuristics depict similar behavior for minimum memory capacity $m_c$  and  increasing the memory capacity slightly does not change the performance of all heuristics. As memory capacity increased further, dynamic variants of heuristics starts performing better and for the moderate memory capacities (close to $2m_c$ ), static order with dynamic correction variants outperform others. \textit{Gilmore-Gomory} heuristic does not achieve good performance. It can be understood from the fact that the task sequence is obtained considering no extra memory is available, but here we apply the same sequence for a different scenario,  overall memory consumed at each point is limited to memory capacity.

@@

	CCSD application operates on tasks of different sizes, hence different heuristics depict distinct behaviors even at minimum memory capacity $m_c$ . Heterogeneity favors to dynamic strategies, therefore both dynamic and static order with dynamic correction based strategies perform good than  static based strategies. Similar to HF,  static order with dynamic correction based strategies outperform others as memory capacity becomes moderate.

@@

	Figure~\ref{fig:ratio_to_optimal_best_ccsd} shows that best variants of dynamic and static order with dynamic correction  strategies achieve similar performance at minimum memory capacity $m_c$. But as memory capacity increases heterogeneity allows static+dynamic strategies to take advantage of static knowledge to get maximum overlap and dynamic correction to select another task in case of memory capacity limitation. Static strategies also start performing better at the end, which indicates that this application has potential for significant communication-computation overlap and pure dynamic strategies are unable to take scheduling decision based on this information.

@@

	\subsection{MILP Based Strategy}
	We use GLPK solver v4.65 to solve the MILP formulation described in Section~\ref{subsec:linearprogrammingformulation}. The solver was unable to handle number of variables required to solve MILP at the scale of our interest. Hence, we solve the linear program iteratively for a small subset of size $k=3,4,5$ (solver was unable to handle number of variables required for $k\ge6$ ) and represent the makespan calculated by this heuristic as $lp.k$. At the boundary of two iterations we fixed the event (communication or computation) of an unfinished task who has  started before the boundary point and consider other events flexible. We compute various  $lp.k$ values for different memory capacities and observe that most of the other heuristics perform better than this heuristic. Figure~\ref{fig:iterativeLpSolution} shows the comparison of different heuristics with MILP based heuristic for various memory capacities of a single trace file.
	\todo[inline]{SKumar: Obtain MILP based makespans for new memory capacities}
	\begin{figure}[htb]
		\includegraphics[scale=0.5]{./results/makespan_with_lp.pdf}
		\caption{Comparision of different heuristics with MILP solution based heuristic.}
		\label{fig:iterativeLpSolution}
	\end{figure}

@@

	In this article, we consider the problem of deciding the order of data transfers between two memory nodes such that overlap of communication and  computation is maximized. With Exascale computing, applications face bottlenecks due to communications. Hence, it is extremely important to achieve the maximum overlap of computation and communication in order to exploit the full potential of the system. We show that determining the order of data transfers is an NP complete problem. We proposed multiple data transfer heuristics and evaluated them on two molecular chemistry kernels, HF and CCSD. Our results show that some of our heuristics achieve significant overlap and perform very close to the lower bound of makespan. We plan to evaluate our strategies on different applications coming from multiple domains. We also plan to study the behavior of our strategies in the context of overlapping CPU-GPU communications with computation. A runtime system aims at exposing different heuristics, to maximize the communication-computation overlap, at developer level and automatically selecting the best one is currently underway.

>>>>

	%%\begin{CCSXML}
	%%	<ccs2012>
	%%	<concept>
	%%	<concept_id>10010520.10010553.10010562</concept_id>
	%%	<concept_desc>Computer systems organization~Embedded systems</concept_desc>
	%%	<concept_significance>500</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10010520.10010575.10010755</concept_id>
	%%	<concept_desc>Computer systems organization~Redundancy</concept_desc>
	%%	<concept_significance>300</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10010520.10010553.10010554</concept_id>
	%%	<concept_desc>Computer systems organization~Robotics</concept_desc>
	%%	<concept_significance>100</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10003033.10003083.10003095</concept_id>
	%%	<concept_desc>Networks~Network reliability</concept_desc>
	%%	<concept_significance>100</concept_significance>
	%%	</concept>
	%%	</ccs2012>
	%%\end{CCSXML}
	\ccsdesc[100]{Computer systems organization~Embedded systems}
	\ccsdesc[100]{Computer systems organization~Redundancy}
	\ccsdesc{Computer systems organization~Robotics}
	\ccsdesc[100]{Networks~Network reliability}

@@

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges come into picture. First, every architecture provides its own efficacy and interface. Therefore, a steep learning curve is required for programmers to take good utilization of all resources. Second, scheduling is a well known NP-Complete optimization problem, and hybrid and distributed resources make this problem harder (we refer ~\cite{webpagescheduling} for a survey on the complexity of scheduling problems and ~\cite{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses, parallel resources, it is hard to obtain a precise model based on prediction of computation and communication timings. Fourth, number of architectures has increased drastically in recent years, therefore it is almost impossible to develop hand tune optimized code for all these architectures. All these observations led to the development of different task based runtime systems. Among several runtimes, we may cite QUARK~\cite{YarKhan:2011:Quark:Manual} and PaRSEC~\cite{parsec} from from ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} form INRIA Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAPPI~\cite{kaapi} from INRIA Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information such as distance to exit (last) node as a priority or affinity of computation towards resources is also provided along with the DAG. Then runtime is responsible for managing scheduling of computations and communications, data transfers among different memories, computation-communication overlap, and load balance.

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities, based on the task graph structure, with computations, which runtime uses to make the execution efficient. In irregular applications, programmers do not know the precise structure of the tasks graphs in advance, tasks are added recursively based on certain sentinel constraints. For such applications, runtime sees a set of independent ready tasks and schedule them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations such that overlap between computations and communications can be maximised. This is the main topic of this article. We prove that order of communications on two memory nodes  with the objective of minimizing the makespan is  NP-Complete. Our proof is inspired from work by ~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem. The main difference with their and our approach is that they consider unit buffer requirement on 1st machine after job completion. On the contrary, our approach is designed where there is not any restriction on requested input data and memory is acquired before starting the data transfer on communication resource. We propose different strategies for runtimes such that it can maximize the overlap of computation and communication for kernels running on two different memory nodes. We also prove that one of our proposed strategies is an optimal one when we do not have any restriction on the capacity of memory node.  We evaluate our strategies on a cluster of homogeneous nodes. But our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper.

@@

	There is also a numerous study pertaining to the scheduling with limited memory and storage, since the work of register allocation for arithmetic expressions by Sethi and Ulman~\cite{Sethi:1970:GOC:321607.321620}.  Sarkar et. al  worked on the scheduling of graphs of smaller-grain tasks with limited memory, where each task requires homogeneous data size ~\cite{vsarkar-pact}. The same work has been extend by Marchal et. al for task graphs where memory requirement of each task is highly irregular~\cite{loris-ipdps18}.

@@

\noindent In order to prove the optimality of makespan for the schedule obtained by the  Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}, we need the following lemma.

@@

			\noindent $\scomp(A) = max(t_1 + CM_A, t_2)$ \\
			\noindent $\scomp(B) = max(\scomp(A) + CP_A,  t_1 + CM_A + CM_B)$\\

%%		\begin{eqnarray*}
%%			\scomp(A) & = & max(t_1 + CM_A, t_2) \\
%%			\scomp(B) & = & max(\scomp(A) + CP_A,  t_1 + CM_A + CM_B)
%%		\end{eqnarray*}

		\noindent Completion time of $B$ in original Schedule = $\scomm(B) + CP_B$\\
		\noindent Completion time of $B$ in swapped Schedule,  $\scomp '(B)  =  max(t_1 + CM_B, t_2)$\\
		\noindent Completion time of $A$ in swapped Schedule, $ \scomp '(A) = max(\scomp'(B) + CP_B, t_1 + CM_B + CM_A)$

%%		\begin{eqnarray*}
%%			\text{Completion time of $B$ in swapped Schedule, } \scomp '(B) & = & max(t_1 + CM_B, t_2)\\
%%			\text{Completion time of $A$ in swapped Schedule, } \scomp '(A) & = & max(\scomp'(B) + CP_B, t_1 + CM_B + CM_A)
%%		\end{eqnarray*}

@@

		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(\scomp(A) + CP_A + CP_B,  t_1 + CM_A + CM_B + CP_B)\\
			& = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B) 
		\end{eqnarray*}
		Case I: $CP_A \ge CM_A, CP_B \ge CM_B, CM_A \le CM_B$
		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B)\\
			& \le& max(t_1 + CM_A + CP_B, t_2  + CP_B, t_1  + CM_B + CP_B) + CP_A\\
			& = & max( t_1  + CM_B + CP_B, t_2  + CP_B) + CP_A \\
			& = & max(max(t_1 + CM_B, t_2) + CP_B, t_1 + CM_A + CM_B) + CP_A\\
			& = & max(\scomp'(B) + CP_B, t_1 + CM_A + CM_B) + CP_A \\
			& = & \scomp'(A) + CP_A
		\end{eqnarray*}
		Case II: $CP_A < CM_A, CP_B < CM_B, CP_A \ge CP_B$
		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B)\\
			& \le & max(t_1 + CM_A + CP_B, t_2+CP_B, t_1 + CM_A + CM_B) + CP_A\\
			& = & max(t_1 + CM_A + CM_B, t_2+ CP_B) + CP_A \\
			& = & max(t_1 + CM_B + CP_B, t_2 + CP_B, t_1 + CM(A) + CM_B) + CP_A \\
			& = & max(max(t_1 + CM_B, t_2) + CP_B, t_1 + CM_A + CM_B) + CP_A \\
			& = & max(\scomp'(B) + CP_B, t_1 + CM_A +CM_B) + CP_A \\
			& = & \scomp'(A) + CP_A
		\end{eqnarray*}
		Case III: $CP_A \ge CM_A, CP_B < CM_B$
		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B)\\
			& \le & max(t_1 + CM_A + CP_B, t_2 + CP_B, t_1 + CM_B + CP_B) + CP_A\\
			& \le & max(t_1 + CM_A + CM_B, t_2 + CP_B, t_1 + CM_B + CP_B) + CP_A \\
			& = & max(max(t_1 + CM_B, t_2) + CP_B, t_1 + CM_A + CM_B) + CP_A \\
			& = & max(\scomp'(B) + CP_B, t_1 + CM_A +CM_B) + CP_A \\
			& = & \scomp'(A) + CP_A
		\end{eqnarray*}
	\end{proof}

@@

		Let the final swapped schedule is $S_{final}$. From Lemma~\ref{lemma:swappingOfTasks}, $makespan(S_{final})$ $ \ge $ $makesapn(S)$. From the construction, $makespan(S_{final})$ $ \le$ $ makesapn(O)$. As $O$ is an optimal schedule, hence $makespan(S) = makesapn(O)$. This completes our proof.
	\end{proof}

@@

		Conversely, if such a partition of $A$ is achievable, then we can construct a feasible schedule $S$ 
		without idle times by the pattern depicted in Fig~\ref{fig:firstSegment}. We have shown that $3PAR$ 
		problem reduces to $DT$, and hence problem $DT$ is NP-complete.

@@

@@

	Fig~\ref{fig:staticOrderExample} shows schedules for all proposed heuristics of this class for the task set of Table~\ref{tab:staticOrderExample}.
	\subsection{Dynamic Selection}
	%%%%string alg_name="largest_communication\_task\_respects\_memory\_restriction";
	%%%%string alg_name="smallest_communication\_task\_respects\_memory\_restriction";
	%%%%string alg_name="maximum_accelerated\_task\_respects\_memory\_restriction";

@@

	Fig~\ref{fig:dynamicSelectionExample} shows schedules for all proposed heuristics of this class for the task set of Table~\ref{tab:dynamicSelectionExample}.

@@

	Fig~\ref{fig:staticOrderDynamicCorrectionsExample} shows schedules for all proposed heuristics of this class for the task set of Table~\ref{tab:staticOrderDynamicCorrectionsExample}.

@@

We use GLPK solver v4.65 to solve the above formulation. The solver was unable to handle number of variables required to solve MILP at the scale of our interest. Hence, we solve the linear program iteratively for a small subset of size $k=3,4,5$ (solver was unable to handle number of variables required for $k\ge6$ ) and represent the makespan calculated by this heuristic as $lp.k$. At the boundary of two iterations we fixed the event (communication or computation) of an unfinished task who has  started before the boundary point and consider other events flexible. We compute various  $lp.k$ values for different memory capacities and observe that most of the other heuristics perform better than this heuristic. Hence, we do not include  this heuristic for the comparison in Section~{\ref{sec:exp}. Figure~\ref{fig:iterativeLpSolution} shows the performance of different heuristics with MILP based heuristics for various memory capacities of a single trace file.
%%	\subsection{MILP Based Strategy}
%%We use GLPK solver v4.65 to solve the MILP formulation described in Section~\ref{subsec:linearprogrammingformulation}. The solver was unable to handle number of variables required to solve MILP at the scale of our interest. Hence, we solve the linear program iteratively for a small subset of size $k=3,4,5$ (solver was unable to handle number of variables required for $k\ge6$ ) and represent the makespan calculated by this heuristic as $lp.k$. At the boundary of two iterations we fixed the event (communication or computation) of an unfinished task who has  started before the boundary point and consider other events flexible. We compute various  $lp.k$ values for different memory capacities and observe that most of the other heuristics perform better than this heuristic. Figure~\ref{fig:iterativeLpSolution} shows the comparison of different heuristics with MILP based heuristic for various memory capacities of a single trace file.
\todo[inline]{SKumar: Obtain MILP based makespans for new memory capacities}
\begin{figure}[htb]
	\includegraphics[scale=0.5]{./results/makespan_with_lp.pdf}
	\caption{Comparision of different heuristics with MILP solution based heuristic.}
	\label{fig:iterativeLpSolution}
\end{figure}

@@

	We refer ~\cite{Gilmore-Gomory:1964} for the detailed algorithm and ~\cite{gitworkrepo} for our implementation. This algorithm does not take memory capacity into consideration and only provides the sequence of processing. We applied the same sequence with different memory capacity restrictions for our experiments and call this heuristic as \textit{Gilmore-Gomory} ($GG$).

	%%	\todo[inline]{may convert 10 subplots to 8 in all performance diagrams}
	\todo[inline]{increase font size in all plots, may be free scale-y}
%%	\todo[inline]{SKumar: decide metrics: ratio to optimal, percentage overlap, 100 - percentage non overlap}

@@

	Figures ~\ref{fig:ratio_to_optimal_hf} and ~\ref{fig:ratio_to_optimal_ccsd} depict the distribution of the performance  of each heuristic for all considered memory capacities, where plots are categorized by memory capacities. For each memory capacity and each heuristic, the box on the plot displays the median, first and last quartile, and the whiskers indicate minimum and maximum values, with outliers are shown by black dots.
	\subsection{HF Performance}	
	\begin{figure*}[htb]
		\includegraphics[scale=0.25]{./results/plots/ratio_to_optimal_selected_hf.pdf}
		\caption{Comparison of different heuristics for HF.}
		\label{fig:ratio_to_optimal_hf}
	\end{figure*}
	As indicated above, HF tasks are highly homogeneous, this is also noticeable in Figure~\ref{fig:ratio_to_optimal_hf}. All heuristics depict similar behavior for minimum memory capacity $m_c$  and  increasing the memory capacity slightly does not change the performance of all heuristics. As memory capacity increased further, dynamic variants of heuristics starts performing better and for the moderate memory capacities (close to $2m_c$ ), static order with dynamic correction variants outperform others. \textit{Gilmore-Gomory} heuristic does not achieve good performance. It can be explained from the fact that the task sequence is obtained considering no extra memory is available, but here we apply the same sequence for a different scenario,  overall memory consumed at each point is limited to memory capacity.

@@

	CCSD application operates on tasks of different sizes, hence different heuristics depict distinct behaviors even at minimum memory capacity $m_c$ . Heterogeneity favors to dynamic strategies, therefore both dynamic and static order with dynamic correction based strategies perform better than  static based strategies. Similar to HF,  static order with dynamic corrections based strategies outperform others as memory capacity becomes moderate.

@@

	Figure~\ref{fig:ratio_to_optimal_best_ccsd} shows that best variants of dynamic and static order with dynamic correction  strategies achieve similar performance at minimum memory capacity $m_c$. But as memory capacity increases, heterogeneity allows static order with dynamic corrections based strategies to take advantage of static knowledge to get maximum overlap and dynamic correction to select another task in case of memory capacity limitation. Static strategies also start performing better at the end, which indicates that this application has potential for significant communication-computation overlap and pure dynamic strategies are unable to take this information into account while making scheduling decisions.

@@

@@

	In this article, we consider the problem of deciding the order of data transfers between two memory nodes such that overlap of communication and  computation is maximized. With Exascale computing, applications face bottlenecks due to communications. Hence, it is extremely important to achieve the maximum overlap of computation and communication in order to exploit the full potential of the system. We show that determining the order of data transfers is an NP complete problem. We proposed several data transfer heuristics and evaluated them on two molecular chemistry kernels, HF and CCSD. Our results show that some of our heuristics achieve significant overlap and perform very close to the lower bound of makespan. We plan to evaluate our strategies on different applications coming from multiple domains. We also plan to study the behavior of our strategies in the context of overlapping CPU-GPU communications with computations. A runtime system aims at exposing different heuristics, to maximize the communication-computation overlap, at developer level and automatically selecting the best one is currently underway.
